---
title: "Bayesian Statistics Presentation"
logo: asu-logo.png
author: "Grina Hwang"
date: "2023-11-07"
output: ioslides_presentation
---


<style>
.gdbar img {
width: 150px !important;
height: 80px !important;
margin: 8px 8px;
}

.gdbar {
width: 200px !important;
height: 100px !important;
}

</style>

<style type="text/css">
body, p, div, h1, h2, h3, h4, h5, ul {
color.black;
font-family: Modern Computer Roman;
}
h2 {
color: #8C1D40;
}

</style>

<style type="text/css">
slides > slide {
background: linear-gradient(#ffffff, #ffffff 85%, #ffffff);
background-color: white;
}
</style>

```{r setup, include=FALSE} 
knitr::opts_chunk$set(echo = TRUE, comment = "", cache=F) 
library(ggplot2)
library(datasets)
library(plotly)
data(USArrests)
```

## What is Bayesian Statistics?

Bayesian statistics is a probabilistic approach that combines prior beliefs with new data to quantify 
uncertainty. It's used for parameter estimation, decision-making under uncertainty, and in various 
fields like machine learning and Bayesian networks, where modeling complex, uncertain relationships is 
crucial. Bayesian statistics is a versatile tool for managing uncertainty and making informed decisions.

## Bayes' Theorem
  
  The fundamental equation of Bayesian statistics is Bayes' theorem:

\[
P(A|B) = \frac{P(B|A) \cdot P(A)}{P(B)}
\]

Where: <br>
P(A|B) is the probability of the hypothesis being true, given the evidence. <br>
P(B|A) is the probability of the evidence occurring if the hypothesis is true. <br>
P(A) is the prior probability of the hypothesis being true. <br>
P(B) is the total probability of the evidence. <br>


## Dataset `US Arrests`
```{r USArrests, echo = TRUE}
data(USArrests)
head(USArrests)
```


## US Arrests for Murder

<font size="4">Before employing Bayesian statistics, it's beneficial to create data visualizations to explore 
and gain insights into the various types of US arrests across different states. This approach allows for a comprehensive examination of the information pertaining to US arrest statistics and their regional distribution. </font>

```{r, echo = F} 
data(USArrests)
murder_summary <- data.frame(State = rownames(USArrests), Murder = USArrests$Murder)

murder_summary <- murder_summary[order(-murder_summary$Murder), ]

ggplot(data = murder_summary, aes(x = reorder(State, -Murder), y = Murder)) +
  geom_bar(stat = "identity", fill = "light pink") +
  labs(title = "Murder Instances by State in USArrests Dataset",
       x = "State",
       y = "Number of Murder Instances") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## US Arrests for Assault

```{r, echo = F} 
murder_summary <- data.frame(State = rownames(USArrests), Murder = USArrests$Murder)
murder_summary <- murder_summary[order(-murder_summary$Murder), ]

assault_summary <- data.frame(State = rownames(USArrests), Assault = USArrests$Assault)
assault_summary <- assault_summary[order(-assault_summary$Assault), ]

library(ggplot2)

ggplot(data = assault_summary, aes(x = reorder(State, -Assault), y = Assault)) +
  geom_bar(stat = "identity", fill = "lightpink") +
  labs(title = "Assault Instances by State",
       x = "State",
       y = "Number of Assault Instances") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Comparison of UrbanPop against Murder, Rape, and Assault

<font size="4">Higher UrbanPop correlates positively with Murder, Assault, and Rape, indicating that more urbanized areas typically exhibit increased crime rates. </font>
```{r, echo = F} 
plot_ly(USArrests, x = ~Murder, y = ~Assault, z = ~UrbanPop, type = "scatter3d", mode = "markers", color = I("pink"))
```


## Using Bayes Theorem to predict future murders in US States

<font size="5">By using Bayes' theorem, future murders can be predicted by utilizing the "USArrests" dataset. This will require installing "rstan" or "brms" packages. Stan is a programming language for Bayesian data analysis.</font>

<font size="3">\[ P(\text{Murder}|\text{UrbanPop, Assault, Rape}) = 
\frac{P(\text{UrbanPop, Assault, Rape}|\text{Murder}) \cdot P(\text{Murder})}{P(\text{UrbanPop, Assault, Rape})} \]
</font>


## Code for the Bayes Theorem to plot the USArrests

```{r}
data(USArrests)
murderdata <- USArrests$Murder
urbanpopdata <- USArrests$UrbanPop
assaultdata <- USArrests$Assault
rapedata <- USArrests$Rape

bayestheorem <- sum(murderdata == "Murder" & urbanpopdata * 
assaultdata * rapedata) / sum(urbanpopdata * assaultdata * rapedata)
```


## Plot for the Bayes Theorem

```{r}
hist(bayestheorem, main = "Histogram of the Bayes Theorem", 
xlab = "P(Murder | UrbanPop, Assault, Rape)",  col = "pink")
```

